---
output: github_document
---


# Welcome to the ERVEBOCOSTS repository

This repository is dedicated to ...


## Getting started

You initially need to clone the repository:
```{bash eval = FALSE}
git clone git@github.com:thibautjombart/ervebocosts.git
cd ervebocosts
```

If you have done so already, make sure you use the latest version of *main* 
branch by typing, in a terminal opened in the root folder:
```{bash eval = FALSE}
git checkout main
git pull
```

The infrastructure is compatible with **R** version 4.5.2.
Initially, you will need to ensure all required packages are installed on your 
computer. To ensure this the first time you use the infrastructure, use:

```{r eval = FALSE}
if (!require(here)) install.packages("here")
if (!require(rio)) install.packages("rio")
if (!require(tibble)) install.packages("tibble")
if (!require(magrittr)) install.packages("magrittr")
```


## Content of the repository

The main files and folders of the repository include:

- `README.Rmd`: the *rmarkdown* file generating the `README.md`
- `README.md`: this file, serving as main documentation for the repository
- `data/`: folder containing the cost data
- `R/`: folder containing R helper functions to exploit the data



## Loading helper functions

To load helper functions, use:

```{r }
if (!require(devtools)) {
  install.packages("devtools")
}
devtools::install_deps(dependencies = TRUE, upgrade = "never")
devtools::load_all()
library(dplyr)
library(tibble)
library(tidyr)
```



## Accessing cost data

Cost data are stored as an export of a REDCap database, which is awkward to use
as-is. We provide a series of functions for importing and processing the data
in more usable formats. First, we import the data using the helper function
`import_data()`

```{r}
raw_data <- import_data()
head(raw_data)
```

The content of the original database is outlined in the file `database_content.md`.


## Getting cost items

The main helper function extracts (where available) data on the outbreaks and 
documented cost items:

```{r}
x <- get_cost_items(raw_data)
x <- as.data.frame(x)
head(x)
```

To subset data, we can use `dply::filter`. For instance, we can get data on 
Ebola-specific costs for IPC since 2018 using:
```{r}
x %>% 
  filter(pathogen == "Ebola", 
         year >= 2018, 
         cost_category == "ipc"
         )

```

The list of documented cost categories in the database is:
```{r}
pull(x, "cost_category") %>% unique() %>% sort()
```






