---
output: github_document
---


# Welcome to the ERVEBOCOSTS repository

This repository is dedicated to ...


## Getting started

You initially need to clone the repository:
```{bash eval = FALSE}
git clone git@github.com:thibautjombart/ervebocosts.git
cd ervebocosts
```

If you have done so already, make sure you use the latest version of *main* 
branch by typing, in a terminal opened in the root folder:
```{bash eval = FALSE}
git checkout main
git pull
```

The infrastructure is compatible with **R** version 4.5.2.
Initially, you will need to ensure all required packages are installed on your 
computer. To ensure this the first time you use the infrastructure, use:

```{r eval = FALSE}
if (!require(here)) install.packages("here")
if (!require(rio)) install.packages("rio")
if (!require(tibble)) install.packages("tibble")
if (!require(magrittr)) install.packages("magrittr")
```


## Content of the repository

The main files and folders of the repository include:

- `README.Rmd`: the *rmarkdown* file generating the `README.md`
- `README.md`: this file, serving as main documentation for the repository
- `data/`: folder containing the cost data



## Accessing cost data

We import the data into the R session and convert it to a `tibble` for easier
visualization:

```{r}
file_path <- here::here("data", "ebola_cost_data.xlsx")
x <- rio::import(file_path) %>% 
  tibble()
names(x) <- epitrix::clean_labels(names(x))
x
```

List of main accessible information:

Outbreak data

- pathogen: column `pathogen`

- location the data was collected for: look for columns named `outbreak_location_country_outbreak_1_choice_[country]` that has a value 'checked'; 
check other location info in "outbreak_location_province_state_region_outbreak_1" and
"outbreak_location_village_municipality_city_outbreak_1" 

- when the outbreak started: columns `grep("outbreak_start", names(x), value = TRUE)`

- when the outbreak stopped: columns `grep("outbreak_end", names(x), value = TRUE)`

- number of confirmed cases: column "number_of_confirmed_cases_outbreak_1"

- number of deaths: column "number_of_deaths_outbreak_1" 

- number of SDBs: column "number_of_safe_burials_outbreak_1"  

- number of vaccinated: column "number_of_vaccines_administered_outbreak_1"

- number of tests: column "number_of_tests_carried_out_outbreak_1"

- number of ETU admissions: column "number_of_etu_admissions_outbreak_1"  

- number of contacts traced: column "number_contacts_traced_outbreak_1"  

Cost data:

- cost category: columns `grep("cost_category", names(x), value = TRUE)`

- cost estimate mean: columns `grep("cost_estimate_or_mean_cost_", names(x), value = TRUE)`

- cost estimate lower bound: columns `grep("cost_range_lower_bound_", names(x), value = TRUE)`

- cost estimate upper bound: columns `grep("cost_range_upper_bound_", names(x), value = TRUE)`

- currency: columns `grep("currency_cost", names(x), value = TRUE)`

