---
output: github_document
---



```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  warning = FALSE, 
  message = FALSE
)
```

# Welcome to the ERVEBOCOSTS repository

This repository is dedicated to extracting Ebola intervention cost data from a REDCap database instance (exported as an `xlsx` file). Cost data are stored in an awkward format, spread across varying numbers of columns, so this infrastructure provides helper functions for extracting information in a more user-friendly way.


## Getting started

You initially need to clone the repository:
```{bash eval = FALSE}
git clone git@github.com:thibautjombart/ervebocosts.git
cd ervebocosts
```

If you have done so already, make sure you use the latest version of *main* 
branch by typing, in a terminal opened in the root folder:
```{bash eval = FALSE}
git checkout main
git pull
```

The infrastructure is compatible with **R** version 4.5.2.
Helper functions are distributed as a dedicated package, so you should not have
to care about dependencies: these will be handled for you (see "Loading helper functions" below).


## Content of the repository

The main files and folders of the repository include:

- `README.Rmd`: the *rmarkdown* file generating the `README.md`
- `README.md`: this file, serving as main documentation for the repository
- `data/`: folder containing the cost data
- `R/`: folder containing R helper functions to exploit the data



## Loading helper functions

To load helper functions, and install all required dependencies, use:

```{r }
if (!require(devtools)) {
  install.packages("devtools")
}
devtools::install_deps(dependencies = TRUE, upgrade = "never")
devtools::load_all()
library(dplyr)
library(tibble)
library(tidyr)
```



## Accessing cost data

Cost data are stored as an export of a REDCap database, which is awkward to use
as-is. We provide a series of functions for importing and processing the data
in more usable formats. First, we import the data using the helper function
`import_data()`

```{r results = "hide"}
raw_data <- import_data()
```
```{r}
head(raw_data)
```

The content of the original database is outlined in the file `database_content.md`.


## Extracting cost items

The main helper function extracts (where available) data on the outbreaks and 
documented cost items:

```{r results = "hide"}
x <- get_cost_items(raw_data)
x <- as.data.frame(x)
```

```{r }
dim(x)
head(x)
```

This `data.frame` contains `r nrow(x)` cost items (rows) broken down as:

- `record_id`: unique ID of the record in the original REDCap database
- `pathogen`: name of the pathogen the cost was recorded for
- `locations`: all locations of the recorded costs
- `year`: year the cost item was recorded
- `start_data`: if the cost item was recorded during an outbreak, the starting date of the outbreak
- `end_data`: if the cost item was recorded during an outbreak, the end date of the outbreak
- `cases`: if the cost item was recorded during an outbreak, the number of confirmed cases
- `deaths`: if the cost item was recorded during an outbreak, the number of deaths
- `sdbs`: if the cost item was recorded during an outbreak, the number of safe and dignified burials (SDBs)
- `vaccination`: if the cost item was recorded during an outbreak, the number of vaccine doses administered
- `tests`: if the cost item was recorded during an outbreak, the number of tests performed
- `admissions`: if the cost item was recorded during an outbreak, the number of admissions to healthcare facilities
- `contact_traced`: if the cost item was recorded during an outbreak, the number of contacts traced
- `cost_category`: the type of activity the cost corresponds to
- `cost_category`: finer-grain precisions on the activity the cost corresponds to
- `cost_summary`: a text-summary of the cost estimates and confidence intervals, if available, including currency used
- `cost_estimate`: the numerical value of the central estimate of the cost 
- `cost_lower`: the numerical value of the lower bound of the cost
- `cost_upper`: the numerical value of the upper bound of the cost

To subset data, we can use `dply::filter`. For instance, we can get data on 
Ebola-specific costs for IPC since 2018 using:
```{r}
x %>% 
  filter(pathogen == "Ebola", 
         year >= 2018, 
         cost_category == "ipc"
         )

```

The list of documented cost categories in the database is:
```{r}
pull(x, "cost_category") %>% unique() %>% sort()
```


The list of documented cost subcategories in the database is:
```{r}
pull(x, "cost_subcategory") %>% unique() %>% sort()
```






